---
title: "covid 19"
author: "Sierra Green"
date: "09/29/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```  



The first chart shows the frequency between female and male who are alive and the frequency from passing. The males and the females frequency of being alive are a lot greater than those who are deceased. The second chart demonstrates the percentage between female/males that are and those who have passed. It shows that females are more capable of staying alive rather then males.

  
## Numerical results


```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This table represents that 98.7% of females have survived covid, leaving 1.4% of them dead. For the males, it leaves at 97% of surving covid, leaving 3% of them dead.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The chi sq test provides the exact numbers. It shows that 1831 were F and 1411 M that were alive. It also shows that 25 F and 44 M died.
The fisher test gives us the odds of dying for males it 44/1411 and the odds of the female dying is 25/1831. The odds ratio is 44/1411 divided by 25/1831 is 2.3. The odds ratio shows that males are 2.3x more likely to die than females.


## Conclusions

